Synthesizing Verified Components for Cyber Assured Systems Engineering \thanks{{DISTRIBUTION STATEMENT A. Approved for public release.} }

### 2021-05-15 paper/agree.tex
add connection between stepFn and AGREE semantics; some other minor tweakage

\subsection{Brief Semantics Definition}

Here the formal semantics of the AGREE contract specification are briefly
presented to make clear the meaning of a high-assurance component. 
These semantics are used in the next section to argue that the synthesis
preserves the same input to output behavior as the AGREE specification.

@@

\begin{comment}
[\emph{So basically $\theta$ covers the inputs and the state, i.e. we can
generate a contig type for the inputs plus state vars.} ]

[Question: what is a \emph{total} environment? One that supplies at
least the bindings in $\theta$ needed to evaluate the spec?]
\end{comment}

The function $\konst{eval}\; s\; \pi$ is defined such that it evaluates $s$ on
the stream, $\pi$, and returns \konst{true} if $s$ is invariant along the
entire stream and \konst{false} otherwise. A guarantee in $s$ is invariant if
its associated expression is true in the context of every step of the stream, while
an eq-statement in $s$ is invariant if its binding in the context of every step
is equivalent to the computed value of its associated expression in that same
context. All guarantees and eq-statements must be invariant in the stream for
the function to return \konst{true}.

The meaning of an AGREE contract
specification is now defined as the set of environment streams on which it is invariant.
%
\[
  \mathcal{L}(s) = \set{\pi \in \Theta_s^* \mid \konst{eval}\ s\ \pi = \konst{true}}
\]
%
\noindent Intuitively, any stream $\pi \in \mathcal{L}(s)$, at each step, binds
the L-values in the eq-statements
in a way that is consistent with their associated expressions and the guarantees are all
true in that same step.

We claim that a synthesized high-assurance component preserves the input to output
relationship in the specification $s$ over every stream in $\mathcal{L}(s)$. 
Let $\pi^\prime
= \konst{SynthEval}\ s\ \pi$ denote a function that synthesizes $s$ and then
evaluates that synthesized component on the stream $\pi$ to create a new
stream $\pi^\prime$ containing added output and other bindings. We
say that two streams are equivalent in regards to a
specification, denoted as $\pi =_s \pi^\prime$, if and only if the two streams are the same length and
agree on bindings for the input and output for $s$ at every step of the streams.
We now formally state the correctness claim for synthesis.
%
\[
  \forall \pi \in \mathcal{L}(s)\ ((\konst{SynthEval}\ s\ I_s(\pi)) =_s \pi)\\
\]
%
where $I_s(\pi)$ returns the corresponding stream that only retains bindings
for inputs in each step and nothing else. The claim is that the synthesized component
generates the same output stream as the specification for any stream
belonging to the specification that is restricted to just input bindings at each step.

>>>>

\subsection{Brief Semantics Definition}
\label{agree-semantics}
Here the formal semantics of the AGREE contract specification are briefly
presented to make clear the meaning of a high-assurance component.
These semantics are used in the next section to argue that the synthesis
preserves the same input to output behavior as the AGREE specification.

@@

%% \begin{comment}
%% [\emph{So basically $\theta$ covers the inputs and the state, i.e. we can
%% generate a contig type for the inputs plus state vars.} ]

%% [Question: what is a \emph{total} environment? One that supplies at
%% least the bindings in $\theta$ needed to evaluate the spec?]
%% \end{comment}

The function $\konst{eval}\; s\; \pi$ evaluates $s$ on the stream,
$\pi$, and returns \konst{true} if $s$ is invariant along the entire
stream and \konst{false} otherwise. A guarantee $\mathcal{G}$ in $s$
is \emph{invariant} if $\mathcal{G}$ is true for each prefix of $\pi$,
while an eq-statement in $s$ is invariant if its binding in the
context of every step is equivalent to the computed value of its
associated expression in that same context. All guarantees and
eq-statements must be invariant in the stream for the function to
return \konst{true}.

The meaning of an AGREE contract specification is now defined as the
set of environment streams on which it is invariant.
%
\[
  \Lang{s} = \set{\pi \in \Theta_s^* \mid \konst{eval}\ s\ \pi = \konst{true}}
\]
%
\noindent Intuitively, any stream $\pi \in \mathcal{L}(s)$, at each step, binds
the L-values in the eq-statements in a way that is consistent with
their associated expressions and the guarantees are all true in that
same step.

We claim that a synthesized high-assurance component preserves the
input to output relationship in the specification $s$ over every
stream in $\Lang{s}$.  Let $\pi^\prime = \konst{SynthEval}\;
s\; \pi$ denote a function that synthesizes $s$ and then evaluates that
synthesized component on the stream $\pi$ to create a new stream
$\pi^\prime$ containing added output and other bindings. We say that
two streams are equivalent in regards to a specification, denoted as
$\pi =_s \pi^\prime$, if and only if the two streams are the same
length and agree on bindings for the input and output for $s$ at every
step of the streams.  We now formally state the correctness claim for
synthesis.
%
\[
  \forall \pi \in \Lang{s}.\; (\konst{SynthEval}\; s\; I_s(\pi)) =_s \pi\\
\]
%
where $I_s(\pi)$ returns the corresponding stream that only retains
bindings for inputs in each step and nothing else. The claim is that
the synthesized component generates the same output stream as the
specification for any stream belonging to the specification that is
restricted to just input bindings at each step.

### 2021-05-15 paper/conclusion.tex
add connection between stepFn and AGREE semantics; some other minor tweakage

We are currently in the process of applying BriefCASE to the design of an application using the Collins Common Avionics Architecture System (CAAS)~\cite{caas} on the CH-47F Chinook helicopter as part of the DARPA CASE program.  Other ongoing work includes adding support for uninterpreted functions, mechanizing the synthesis proof in HOL4 and lifting the proof results to infinite streams.

>>>>

We are currently in the process of applying BriefCASE to the design of
an application using the Collins Common Avionics Architecture System
(CAAS)~\cite{caas} on the CH-47F Chinook helicopter as part of the
DARPA CASE program.  Other ongoing work and future work includes
adding support for uninterpreted functions, mechanizing the
correctness-of-synthesis proofs in HOL4, and lifting the proof results
to infinite streams.

### 2021-05-15 paper/synthesis.tex
Added example section reference to Synthesis section
add connection between stepFn and AGREE semantics; some other minor tweakage

The \konst{stepFn} for the monitor of the example is displayed in \figref{fig:monitor-cakeml}.

\newsavebox{\monFn}
\begin{lrbox}{\monFn}
\begin{lstlisting}[style=myML]
stepFn (Request,Response)
       (req,rsp,current,previous,policy,alert) =
let val stateVars' =
     if !initStep then
        let val req = event(Request)
            val rsp = event(Response)
            val current = (req = rsp)
            val previous = req and not(rsp)
            val policy = current or previous
            alert = not policy
        in (req,rsp,current,previous,policy,alert)
        end
     else
        let val req = event(Request)
            val rsp = event(Response)
            val current = (req = rsp)
            val previous = pre(req and not rsp) and (not req and rsp)
            val policy = current or previous
            alert = (is_latched and pre(alert)) or not(policy)
        in (req,rsp,current,previous,policy,alert)
        end
    val (_,rsp',_,_,_,alert') = stateVars'
    val Alert = if alert' then Some () else None
    val Output =
       if alert' then None else
       if rsp'   then Some Response
       else None
in
   (stateVars', (Alert,Output))
end
\end{lstlisting}
\end{lrbox}

@@

Possibly some material to add here ...

>>>>

The \konst{stepFn} for the monitor of the example described in
Section~\ref{sec:example} is displayed in \figref{fig:monitor-cakeml}.

\newsavebox{\monFn}
\begin{lrbox}{\monFn}
\begin{lstlisting}[style=myML]
stepFn (Request,Response)
       (req,rsp,current,previous,policy,alert) =
let val stateVars' =
     if !initStep then
        let val req = event(Request)
            val rsp = event(Response)
            val current = (req = rsp)
            val previous = req and not(rsp)
            val policy = current or previous
            val alert = not policy
            val () = (intStep := False)
        in (req,rsp,current,previous,policy,alert)
        end
     else
        let val req = event(Request)
            val rsp = event(Response)
            val current = (req = rsp)
            val previous = pre(req and not rsp) and (not req and rsp)
            val policy = current or previous
            val alert = (is_latched and pre(alert)) or not(policy)
        in (req,rsp,current,previous,policy,alert)
        end
    val (_,rsp',_,_,_,alert') = stateVars'
    val Alert = if alert' then Some () else None
    val Output =
       if alert' then None else
       if rsp'   then Some Response
       else None
in
   (stateVars', (Alert,Output))
end
\end{lstlisting}
\end{lrbox}

@@

Intutively, for monitor specification $s$, \konst{stepFn} is the
concrete embodiment of $\konst{SynthEval}\;s$, as defined in Section
\ref{agree-semantics}. Its correctness amounts to showing that, given
a sequence of inputs, and an initial state meeting the initialization
constraints, iterating \konst{stepFn} produces a $\pi \in
\Lang{s}$. Morevover, taking the union over all input sequences and
initial states produces $\Lang{s}$ itself.
